---
title: "Title: CSIT 165 Group 4 Project 1"
output: html_document
author: "Name:  Hyun Lee\nPartner: Kiran Jayasinghe"
date: "Date: `r Sys.Date()`"
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
---

### Introduction
The World Health Organization has recently employed a new data science initiative, *CSIT-165*, that uses data science to characterize pandemic diseases.
*CSIT-165* disseminates data driven analyses to global decision makers.
*CSIT-165* is a conglomerate comprised of two fabricated entities: *Global Health Union (GHU)* and *Private Diagnostic Laboratories (PDL)*. 

> [2019 Novel Coronavirus COVID-19 (2019-nCoV) Data Repository by John Hopkins CSSE](https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series)

Data for 2019 Novel Coronavirus is operated by the John Hopkins University Center for Systems Science and Engineering (JHU CSSE).
Data includes daily time series CSV summary tables, including confirmations, recoveries, and deaths. 
Country/region are countries/regions hat conform to World Health Organization (WHO).
Lat and Long refer to coordinates references for the user. 
Date fields are stored in MM/DD/YYYY format.

```{r setup, include = FALSE}
# Loading packages:
library(knitr)
library(kableExtra)
library(magrittr)
library(readr)
library(geosphere)
library(readr)
library(kableExtra) # it was the only package available in RStudio, didn't want to install kable

# Loading data sets from RAW githubusercontent; curl and wget was not available
confirmed_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
death_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"
recovered_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv"

# Download files to local; curl and wget not available
download.file(confirmed_url, basename(confirmed_url))
download.file(death_url, basename(death_url))
download.file(recovered_url, basename(recovered_url))

confirmed_df <- read_csv(url(confirmed_url))
death_df <- read_csv(url(death_url))
recovered_df <- read_csv(url(recovered_url))
```


### Project Objectives

#### Objective 1
> CSIT-165’s first objective is to determine where COVID-19 originated from. Predict where the origin started based on the area with the greatest number of confirmations and deaths on the first recorded day in the data set. Show this is the origin using an if statement.

```{r, include=TRUE, echo=TRUE}
# Identifying the maximum value on the first day
max_conf <- max(confirmed_df$'1/22/20')
max_death <- max(death_df$`1/22/20`)

# Identifying the row of the max column
row_conf <- as.integer(rownames(confirmed_df)[confirmed_df$`1/22/20` == max_conf])
country_conf <- confirmed_df[[row_conf, 2]]
country_origin <- confirmed_df[row_conf, 1:4]

row_death <- as.integer(rownames(death_df)[death_df$`1/22/20` == max_death])
country_death <- death_df[[row_death,2]]

# If condition to identify the most likely origin
if (country_death == country_conf)
{
  cat("Country of origin for COVID19 is most likely ", country_conf)
}
```

#### Objective 2
> Where is the most recent area to have a first confirmed case? To do this, you will need to use a for loop, if statement, and subsets.

```{r, include=TRUE, echo=TRUE}
# Data frame named new_conf that has location with second to last day and lasy day
new_conf <- confirmed_df[,1:4]
new_conf$'secondlast' <- confirmed_df[, ncol(confirmed_df)-1] # add column of second to last day
new_conf$'last' <- confirmed_df[,ncol(confirmed_df)] # add column of last day

# Changing columns of data frame to vectors
last_conf <- new_conf$last
vector_last_conf <- last_conf[[colnames(last_conf)]] # from data frame to vector for binary operations
second_conf <- new_conf$secondlast
vector_second_conf <- second_conf[[colnames(second_conf)]] # from data frame to vector for binary operations
new_conf <- data.frame(confirmed_df[,1:4], vector_second_conf, vector_last_conf, stringsAsFactors = FALSE)

# If the difference between last day and second to last day is greater than zero, it indicates a new set of confirmed cases.
conf_diff <- new_conf$vector_last_conf - new_conf$vector_second_conf
max_conf <- max(conf_diff) # outputs max number
max_diff <- data.frame(confirmed_df[,1:4], conf_diff, stringsAsFactors = FALSE) # makes DF for area identification

# Using loop and if condition to identify the country 
for (i in conf_diff) {
  if (i == max_conf) {
    max_row <- rownames(max_diff)[max_diff$conf_diff == i]
    max_ans <- max_diff[as.numeric(max_row),]
    
    cat("The area with the greatest amount of new confirmations for the past two days is ", max_ans$Country.Region)
}
}
```

#### Objective 3 (completed by Kiran, edited by Hyun)
> How far away are the areas from objective 2 from where the first confirmed case(s) occurred? Please provide answer(s) in terms of miles. Use the function distm from the R package geosphere to calculate the distance between two coordinates in meters (geosphere::distm). You will need to convert the value returned by distm from meters to miles (this conversion is simple and can be found online). Please use a table or printed statement to describe what Province/State and Country/Region first confirmed cases occurred as well as the distance (in miles) away from the origin. Please print the following: {recent region} is {distance in miles} away from {origin city, origin country}.

```{r, include=TRUE, echo=TRUE}
library(geosphere)
Origin = c(country_origin$Long, country_origin$Lat)
Recent = c(max_ans$Long, max_ans$Lat)

distance_between_R_and_O_in_meters = distm(Origin, Recent, fun = distGeo)
distance_between_R_and_O = (distance_between_R_and_O_in_meters)*(0.000621371)

cat("The most recent active country,", max_ans$Country.Region, "is" , distance_between_R_and_O, "meters away from country of origin,", country_origin$`Country/Region`)
```

#### Objective 4 (completed by Kiran, edited by Hyun)
> Which area of the world currently has the lowest risk score (if more than one, display the one with the most confirmations)? Which area of the world currently has the highest risk score (if more than one, display the one with the most confirmations)? How do risk scores in these areas compare to global risk score? Why might it be helpful to calculate metrics like risk scores for different areas of the world and what would their limitations be (what assumptions does risk score make and what important variables might be left out)?

```{r, include=TRUE, echo=TRUE}
# Excluding all cruise ships
no_ship_confirmed_data = confirmed_df[!(confirmed_df$Lat == "NA"|confirmed_df$Lat == 0 ),]
no_ship_deaths_data = death_df[!(death_df$Lat == "NA"|death_df$Lat == 0),]
risk_factor_data = 100*(no_ship_deaths_data[,5:ncol(no_ship_deaths_data)]/no_ship_confirmed_data[,5:ncol(no_ship_confirmed_data)])

# Identifying the region with low risk
min_risk_factor = min(risk_factor_data[ncol(risk_factor_data)], na.rm = TRUE)
areas_w_min_risk_factor = which(risk_factor_data[ncol(risk_factor_data)] == min(risk_factor_data[ncol(risk_factor_data)], na.rm = TRUE))
max_conf_w_min_risk = which(no_ship_confirmed_data[ncol(no_ship_confirmed_data)] == max(no_ship_confirmed_data[areas_w_min_risk_factor,ncol(no_ship_confirmed_data)]))
min_risk_row_data = no_ship_confirmed_data[max_conf_w_min_risk,]
min_risk_prov = min_risk_row_data$`Province/State`
min_risk_country = min_risk_row_data$`Country/Region`

cat("Province,", min_risk_prov, "from", min_risk_country, "has the lowest risk score of", min_risk_factor, "%.")

# Identifying the region with the maximum risk
max_risk_factor = max(risk_factor_data[ncol(risk_factor_data)], na.rm = TRUE)
areas_w_max_risk_factor = which(risk_factor_data[ncol(risk_factor_data)] == max(risk_factor_data[ncol(risk_factor_data)], na.rm = TRUE))
max_conf_w_max_risk = which(no_ship_confirmed_data[ncol(no_ship_confirmed_data)] == max(no_ship_confirmed_data[areas_w_max_risk_factor,ncol(no_ship_confirmed_data)]))
max_risk_row_data = no_ship_confirmed_data[max_conf_w_max_risk,]
max_risk_prov = max_risk_row_data$`Province/State`
max_risk_country = max_risk_row_data$`Country/Region`

cat("Province,", max_risk_prov, "from", max_risk_country, "has the highest risk score with", max_risk_factor, "%.")

# Identifying current global status 
global_current_deaths = sum(death_df[ncol(death_df)])
global_current_conf = sum(confirmed_df[ncol(confirmed_df)])

global_risk_factor = 100*(global_current_deaths)/(global_current_conf)
if(global_risk_factor > min_risk_factor) {
  print('The global risk factor is greater than the area with the minimum risk factor.')
}else
  {
  print('The global risk factor is less than the area with the minimum risk factor.')
}

if(global_risk_factor > max_risk_factor) {
  print('The global risk factor is greater than the area with the maximum risk factor.')
}else
  {
  print('The global risk factor is less than the area with the maximum risk factor.')
}
```
> Risk scores are calculated because it is beneficial to understand the ratio between deaths and confirmations: it's directly correlating the confirmed cases to death, which is not always true. 
Out of the confirmed cases, there are recovered cases. The ratio between recovered and confirmation might add more details to the status of the area. For instance, if an area has high risk with high recovery rate, it might indicate that their preventative care needs work, but their treatment for COVID19 might be exemplary. 

#### Objective 5 (completed by Kiran)
> You are asked to make two tables with the top 5 countries that have the most COVID-19 related confirmations and and deaths. Make sure to include all of the counts for the country, not just the counts for one area in the country. To do this we will need to sum all of the values for each country, create new data frames from these values, and use the package kable to convert those data frames into tables.

```{r, include=TRUE, echo=TRUE}
deaths_by_country = aggregate(death_df[,5:ncol(death_df)],by = list(death_df$`Country/Region`),FUN = sum, na.rm = TRUE)
sum_deaths_by_country = (deaths_by_country[,c(1,ncol(deaths_by_country))])

sorted_sum_deaths_by_country = sum_deaths_by_country[order(sum_deaths_by_country[,2], decreasing = TRUE),]

conf_by_country = aggregate(confirmed_df[,5:ncol(confirmed_df)],by = list(confirmed_df$`Country/Region`),FUN = sum, na.rm = TRUE)
sum_conf_by_country = (conf_by_country[,c(1,ncol(conf_by_country))])

sorted_sum_conf_by_country = sum_conf_by_country[order(sum_conf_by_country[,2], decreasing = TRUE),]

# Make table
kable(sorted_sum_deaths_by_country[1:5,],caption = "Top 5 Countries with Covid-Related Deaths", row.names = FALSE, col.names = c("Country","Total Deaths"))

kable(sorted_sum_conf_by_country[1:5,],caption = "Top 5 Countries with Covid Cases", row.names = FALSE, col.names = c("Country","Total Cases"))
```

#### GitHub repo: [rgroup4_project1](https://github.com/unleel/rgroup4_project1)
### GitHub Log
```{bash gitlog} 
git log --pretty=format:"%nSubject: %s%nAuthor: %aN%nDate: %aD%nBody: %b"
```
